### 2021-05-03 inst/manuscript/supplement.Rmd

---
title: "Supplemental Information"
subtitle: "Hypoxia uncouples HIF gene transcription and metabolic flux in proliferating primary cells"

author:
- Courtney A. Copeland: 
    institute: 1
- Benjamin A. Olenchock: 
    institute: 1, 2
- David R. Ziehr:
    institute: 1, 4
- Sarah McGarrity: 
    institute: 1, 5
- Jamey D. Young: 
    institute: 3
- Joseph Loscalzo: 
    institute: 1
- William M. Oldham: 
    institute: 1
    email: woldham@bwh.harvard.edu
    correspondence: true
institute:
- 1: Department of Medicine, Brigham and Women's Hospital and Harvard Medical School, Boston, MA
- 2: Regeneron Pharmaceuticals, Tarrytown, NY
- 3: Departments of Chemical & Biomolecular Engineering and Molecular Physiology & Biophysics, Vanderbilt University, Nashville, TN
- 4: Department of Medicine, Massachusetts General Hospital and Harvard Medical School, Boston, MA
- 5: Center for Systems Biology, School of Health Sciences, University of Iceland, Reykjavik, Iceland

bibliography_main: library.bib
bibliography_software: packages.bib
csl: cell-metabolism.csl
link-citations: yes

output: 
  # bookdown::pdf_document2:
  #   # template: template.tex
  #   keep_tex: false
  #   df_print: kable
  #   toc: false
  #   number_sections: false
  #   latex_engine: xelatex
  #   pandoc_args:
  #     - '--lua-filter=scholarly-metadata.lua'
  #     - '--lua-filter=author-info-blocks.lua'
  #     - '--lua-filter=pagebreak.lua'
  #     - '--lua-filter=multiple-bibliographies.lua'

  bookdown::word_document2:
    reference_docx: template.docx
    df_print: kable
    number_sections: false
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
      - '--lua-filter=pagebreak.lua'
      - '--lua-filter=multiple-bibliographies.lua'

compact-title: true
geometry: margin = 0.75in
fontsize: 11pt
mainfont: Calibri
graphics: yes
header-includes:
- \usepackage{lineno}
- \linenumbers
- \setlength\linenumbersep{1cm}
- \usepackage{setspace}
- \doublespacing
- \makeatletter\renewcommand*{\fps@figure}{htbp}\makeatother
- \setlength{\parskip}{0.75\baselineskip plus 2pt minus 2pt}
- \usepackage{float}
---

```{r setup, include = FALSE}
# library(Copeland.2021.hypoxia.flux)
devtools::load_all()
library(magrittr)

knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  out.width = "50%", 
  fig.align = "center", 
  fig.pos = "H", 
  out.extra = ""
)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

knitr::opts_knit$set(kable.force.latex = FALSE)

options(tinytex.verbose = TRUE)

# kable function
my_kable <- function(data, ...) {
  kableExtra::kable(data, booktabs = TRUE, linesep = "", ...) %>% 
    kableExtra::kable_styling(latex_options = c("HOLD_position"), font_size = 11)
}

withr::with_dir(here::here(), {
  hyp <- targets::tar_read(lf_hypoxia_table)
  bay <- targets::tar_read(lf_bay_table)
})

```

\newpage

# Tables

<br>

```{r results = "asis"}
hyp %>% 
  flextable::set_caption("Table 1: Lung fibroblast fluxes in 21% and 0.5% oxygen") %>% 
  flextable::flextable_to_rmd()
```

\newpage

```{r results = "asis"}
bay %>% 
  flextable::set_caption("Table 2: Lung fibroblast fluxes following DMSO and BAY treatment") %>% 
  flextable::flextable_to_rmd()
```